---
title: "R Notebook"
output: html_notebook
---

# Example

```{r}
#install.packages("vars") 
#install.packages("astsa") 
#install.packages("mFilter")
#install.packages("tsModel")
#install.packages("tseries")
library(tseries)
library(tsModel)
library(mFilter)
library(vars)
library(astsa)

#creates a vector of response variables
x = cbind(cmort, tempr, part) 

#VAR command does estimation of AR models using ordinary least squares while simultaneously fitting the trend, intercept, and ARIMA model
plot.ts(x , main = "", xlab = "")
fitvar1= VAR(x, p=1, type="both")
summary(fitvar1)
```

```{r}
# Removing NAs
#library(readr)
#hilo <- read_csv("HiloData.csv")

hilo2015.fixed <- na.omit(hilo2015)


# Saving as numeric vectors
riverflow <- as.numeric(hilo2015.fixed$cms)
chlorophyll <- as.numeric(hilo2015.fixed$ChlRFU)
dissolved_o2 <- as.numeric(hilo2015.fixed$Doper)
salinity <- as.numeric(hilo2015.fixed$saltppt)
temp <- as.numeric(hilo2015.fixed$TempC)
turbidity <- as.numeric(hilo2015.fixed$Corr.NTU)

#creates a vector of response variables
h_chlorophyll <-  cbind(riverflow, chlorophyll)
h_dissolvedo2 <- cbind(riverflow, dissolved_o2)
h_salinity <- cbind(riverflow, salinity)
h_temp <- cbind(riverflow, temp)
h_turbidity <- cbind(riverflow, turbidity)

#VAR command does estimation of AR models using ordinary least squares while simultaneously fitting the trend, intercept, and ARIMA model
#plot.ts(h , main = "", xlab = "")

fitvar1_chlorophyll= VAR(h_chlorophyll, p=1, type="both")
summary(fitvar1_chlorophyll)

fitvar1_dissolvedo2= VAR(h_dissolvedo2, p=1, type="both")
summary(fitvar1_dissolvedo2)

fitvar1_salinity= VAR(h_salinity, p=1, type="both")
summary(fitvar1_salinity)

fitvar1_temp= VAR(h_temp, p=1, type="both")
summary(fitvar1_temp)

fitvar1_turbidity= VAR(h_turbidity, p=1, type="both")
summary(fitvar1_turbidity)
```
  

```{r}
# 2nd Lag
fitvar2_chlorophyll= VAR(h_chlorophyll, p=2, type="both")
summary(fitvar2_chlorophyll)

fitvar2_dissolvedo2= VAR(h_dissolvedo2, p=2, type="both")
summary(fitvar2_dissolvedo2)

fitvar2_salinity= VAR(h_salinity, p=2, type="both")
summary(fitvar2_salinity)

fitvar2_temp= VAR(h_temp, p=2, type="both")
summary(fitvar2_temp)

fitvar2_turbidity= VAR(h_turbidity, p=2, type="both")
summary(fitvar2_turbidity)
```


```{r}
# Ask about what this is

acf(residuals(fitvar2_salinity), plot = FALSE)
acf(residuals(fitvar2_salinity),plot = TRUE, ylim=c(0,0.2))

acf(residuals(fitvar2_chlorophyll), plot = FALSE)
acf(residuals(fitvar2_chlorophyll),ylim=c(0,0.2))

acf(residuals(fitvar2_dissolvedo2), plot = FALSE)
acf(residuals(fitvar2_dissolvedo2),ylim=c(0,0.2))

acf(residuals(fitvar2_salinity), plot = FALSE)
acf(residuals(fitvar2_salinity),ylim=c(0,0.2))

acf(residuals(fitvar2_temp), plot = FALSE)
acf(residuals(fitvar2_temp), ylim=c(0,0.2))

acf(residuals(fitvar2_turbidity), plot = FALSE)
acf(residuals(fitvar2_turbidity), ylim=c(0,0.2))

```

```{r}
# 3nd Lag
fitvar3_chlorophyll= VAR(h_chlorophyll, p=3, type="both")
summary(fitvar3_chlorophyll)

fitvar3_dissolvedo2= VAR(h_dissolvedo2, p=3, type="both")
summary(fitvar3_dissolvedo2)

fitvar3_salinity= VAR(h_salinity, p=3, type="both")
summary(fitvar3_salinity)

fitvar3_temp= VAR(h_temp, p=3, type="both")
summary(fitvar3_temp)

fitvar3_turbidity= VAR(h_turbidity, p=3, type="both")
summary(fitvar3_turbidity)
```

```{r}
acf(residuals(fitvar3_salinity), plot = FALSE)
acf(residuals(fitvar3_salinity),plot = TRUE, ylim=c(0,0.2))

acf(residuals(fitvar3_chlorophyll), plot = FALSE)
acf(residuals(fitvar3_chlorophyll),ylim=c(0,0.2))

acf(residuals(fitvar3_dissolvedo2), plot = FALSE)
acf(residuals(fitvar3_dissolvedo2),ylim=c(0,0.2))

acf(residuals(fitvar3_salinity), plot = FALSE)
acf(residuals(fitvar3_salinity),ylim=c(0,0.2))

acf(residuals(fitvar3_temp), plot = FALSE)
acf(residuals(fitvar3_temp), ylim=c(0,0.2))

acf(residuals(fitvar3_turbidity), plot = FALSE)
acf(residuals(fitvar3_turbidity), ylim=c(0,0.2))

```

# MODELING
```{r}
# IDK if we need this
# Autocorrelation of each variables 
# To detect non-randomness in data
# To identify an appropriate time series model if the data are not random
Chl.acf <- acf(chlorophyll, main = "output")
sal.acf <- acf(salinity, main = "output")
turb.acf <- acf(turbidity, main = "output")
temp.acf <- acf(temp, main = "output")
diss.acf <- acf(dissolved_o2, main = "output")
```
# Dickey Fuller Test

```{r}
adf.test(hilo2015.fixed$cms, alternative= "stationary")

adf.test(hilo2015.fixed$ChlRFU, alternative= "stationary")

adf.test(hilo2015.fixed$Doper, alternative= "stationary")

adf.test(hilo2015.fixed$saltppt, alternative= "stationary")

adf.test(hilo2015.fixed$TempC, alternative= "stationary")

adf.test(hilo2015.fixed$Corr.NTU, alternative= "stationary")
```

# Model Selection and estimation
```{r}
chl.select <- VARselect(h_chlorophyll, lag.max=50, type="none", season = 3) #23
chl.select$selection

sal.select <- VARselect(h_salinity, lag.max=50, type="none", season = 3) #21
sal.select$selection

turb.select <- VARselect(h_turbidity, lag.max=50, type="none", season = 3) #11
turb.select$selection

temp.select <- VARselect(h_temp, lag.max=50, type="none", season = 3) #21
temp.select$selection

diss.select <- VARselect(h_dissolvedo2, lag.max=50, type="none", season = 3) #26
diss.select$selection

# returns the number of lags depending on the model selection 
# SC = BIC
# we will use BIC since it does not depend on the number of parameters
```


# Using varselect lag outcomes
```{r}
# varselect lags

fitvar23_chlorophyll= VAR(h_chlorophyll, p=23, type="none")
summary(fitvar3_chlorophyll)

fitvar26_dissolvedo2= VAR(h_dissolvedo2, p=26, type="none")
summary(fitvar3_dissolvedo2)

fitvar21_salinity= VAR(h_salinity, p=21, type="none")
summary(fitvar3_salinity)

fitvar21_temp= VAR(h_temp, p=21, type="none")
summary(fitvar3_temp)

fitvar11_turbidity= VAR(h_turbidity, p=11, type="none")
summary(fitvar3_turbidity)
```

# Using varselect lag outcomes Pt 2
```{r}
# varselect lags

fitvar5_chlorophyll= VAR(h_chlorophyll, p=5, type="none")
summary(fitvar3_chlorophyll)

fitvar23_dissolvedo2= VAR(h_dissolvedo2, p=23, type="none")
summary(fitvar3_dissolvedo2)

fitvar11_salinity= VAR(h_salinity, p=11, type="none")
summary(fitvar3_salinity)

fitvar19_temp= VAR(h_temp, p=19, type="none")
summary(fitvar3_temp)

fitvar6_turbidity= VAR(h_turbidity, p=6, type="none")
summary(fitvar3_turbidity)
```

# To consider the model fit we can perform some diagnostic tests on residuals of the model
# Diagnostic test on residuals 
NULL: there is no serial correlation of any order up to p

```{r}
# We reject the null for all variables ?
# For all variables there is evience of serial correlation
  # If they are serially correlated, they are not independent / not random
turb.serial <- serial.test(fitvar11_turbidity)
turb.serial

sal.serial <- serial.test(fitvar21_salinity, lags.pt = 21)
sal.serial

temp.serial <- serial.test(fitvar21_temp, lags.pt = 21)
temp.serial

chl.serial <- serial.test(fitvar23_chlorophyll, lags.pt = 23)
chl.serial

diss.serial <- serial.test(fitvar26_dissolvedo2 , lags.pt = 26)
diss.serial

```
```{r}
# autocorrelations with plots
# there are some residuals outside of the interval 

acf(residuals(fitvar21_salinity), plot = FALSE)
acf(residuals(fitvar21_salinity),plot = TRUE, ylim=c(0,0.1))

acf(residuals(fitvar23_chlorophyll), plot = FALSE)
acf(residuals(fitvar23_chlorophyll),ylim=c(0,0.1))

acf(residuals(fitvar26_dissolvedo2), plot = FALSE)
acf(residuals(fitvar26_dissolvedo2),ylim=c(0,0.1))

acf(residuals(fitvar21_temp), plot = FALSE)
acf(residuals(fitvar21_temp), ylim=c(0,0.1))

acf(residuals(fitvar11_turbidity), plot = FALSE)
acf(residuals(fitvar11_turbidity), ylim=c(0,0.1))
```

# Test for Heteroscedasticity (oposite of homoscedasticity)
NULL: absence of hetero
ALT: hetero
```{r}
# we reject all :(

turb.arch <- arch.test(fitvar11_turbidity, lags.multi = 11, multivariate.only = TRUE)
turb.arch

sal.arch <- arch.test(fitvar21_salinity, lags.multi = 11, multivariate.only = TRUE)
sal.arch

chl.arch <- arch.test(fitvar23_chlorophyll, lags.multi = 11, multivariate.only = TRUE)
chl.arch

diss.arch <- arch.test(fitvar26_dissolvedo2, lags.multi = 11, multivariate.only = TRUE)
diss.arch

temp.arch <- arch.test(fitvar21_temp, lags.multi = 11, multivariate.only = TRUE)
temp.arch
```

# Normality test
We could use the test of the example or use the histogram
All of the histograms were strongly skewed

# Conclusions
Our model does not meet the majority of the assumptions. Our model is not good enough to predict what will happen but it might be good enough to describe what is going on in our data. We could just use our model to describe how River flow and the other variables behave in our data set

# IRF
```{r}
cmsturb <- irf(fitvar2, impulse = "riverflow", response = "turbidity", n.ahead = 8, ortho = FALSE, runs = 1000)
cmsturb
plot(cmsturb) 
```

```{r}
cmssalin <- irf(fitvar2, impulse = "riverflow", response = "salinity", n.ahead = 8, ortho = FALSE, runs = 1000)
cmssalin
plot(cmssalin)
```


```{r}
cmschlor <- irf(fitvar2, impulse = "riverflow", response = "chlorophyll", n.ahead = 8, ortho = FALSE, runs = 1000)
cmschlor
plot(cmschlor)
```

```{r}
cmsdisso2 <- irf(fitvar2, impulse = "riverflow", response = "dissolved_o2", n.ahead = 8, ortho = FALSE, runs = 1000)
cmsdisso2
plot(cmsdisso2)
```

```{r}
cmstemp <- irf(fitvar2, impulse = "riverflow", response = "temp", n.ahead = 8, ortho = FALSE, runs = 1000)
cmstemp
plot(cmstemp)
```

```{r}
#install.packages("smooth")
library(smooth)
hilo2015$logcms <- NULL
hilo2015$chlcal <- NULL
hilo2015$ChlRF <- NULL
which(is.na(hilo2015))
var.hilo2015 <- hilo2015[-c(which(is.na(hilo2015))),]  
sum(is.na(var.hilo2015))
hilo2015 <- na.omit(hilo2015)
sum(is.na(var.hilo2015))

hilo2015.fixed <- na.omit(hilo2015)
sum(is.na(hilo2015.fixed))
```

 
























