---
title: "2013-2015 Descriptives"
author: "Brianna Cirillo & Odalys Barrientos"
output: html_notebook
---

# NEW DATA SET-Modified Data 2013-2015
```{r}
library(mosaic)
library(ggplot2)
library(readr)

 
hilomodified <- read_csv("~/Documents/MA321-SP21/HiloBay-Adolf/HiloBay_MA321_Spring21/NewDataframes/HiloModified13-15.csv")

# start date: 2013-01-01 00:00:00
# end date: 2015-12-31 23:00:00

head(hilomodified)
tail(hilomodified)

```

# Descriptives on all variables MODIFIED DATA SET
## River Flow MODIFIED
```{r}
length(hilomodified$logcms[which(is.na(hilomodified$logcms)==TRUE)]) # No NAs, YAY!


RiverFlowMod <- ggplot(hilomodified,  aes(x = date, y = cms)) + 
  geom_line()

print(RiverFlowMod + ggtitle("River Flow")+labs(x="Time", y = "River Flow - cubic meters per second"))

favstats(hilomodified$cms)
```
## CHL MODIFIED
```{r}
sum(is.na(hilomodified$ChlRFU)==TRUE) # 1884 NAs

CHLMod <- ggplot(hilomodified,  aes(x = date, y = ChlRFU)) + 
  geom_line()

print(CHLMod + ggtitle("Chlorophyll ")+labs(x="Time", y = "Chlorophyll  - relative fluorescence units (RFU)"))

favstats(hilomodified$ChlRFU)
```
## Turbitity MODIFIED
```{r}
length(hilomodified$Corr.NTU[which(is.na(hilomodified$Corr.NTU)==TRUE)]) # 2704 NAs

TURBMod <- ggplot(hilomodified,  aes(x = date, y = Corr.NTU)) + 
  geom_line()

print(TURBMod + ggtitle("Turbidity ")+labs(x="Time", y = "Turbidity - Nephelometric Turbidity Units (NTU)"))

favstats(hilomodified$Corr.NTU)
```
## Salinity MODIFIED
```{r}
length(hilomodified$saltppt[which(is.na(hilomodified$saltppt)==TRUE)]) # 2267 NAs

SALTMod <- ggplot(hilomodified,  aes(x = date, y = saltppt)) + 
  geom_line()

print(SALTMod + ggtitle("Salinity")+labs(x="Time", y = "Salinity - unit parts per thousand (PPT)"))

favstats(hilomodified$saltppt)
favstats(hilomodified$TempC)
favstats(hilomodified$Doper)
```
## Temperature MODIFIED
```{r}
length(hilomodified$TempC[which(is.na(hilomodified$TempC)==TRUE)]) # 1702 NAs

TempMod <- ggplot(hilomodified,  aes(x = date, y = TempC)) + 
  geom_line()

print(TempMod + ggtitle("Temperature")+labs(x="Time", y = "Temperature - C"))


favstats(hilomodified$TempC)
```

## Dissolved Oxygen MODIFIED
```{r}
length(hilomodified$Doper[which(is.na(hilomodified$Doper)==TRUE)]) # 2267 NAs

DissoMod <- ggplot(hilomodified,  aes(x = date, y = Doper)) + 
  geom_line()

print(DissoMod + ggtitle("Dissolved Oxygen")+labs(x="Time", y = "Dissolved Oxygen"))

favstats(hilomodified$Doper)
```

===================================================
# Histograms MODIFIED 

## River Flow MODIFIED
```{r}
hist(hilomodified$cms, main = "Histogram of River Flow", xlab = "River Flow", breaks =90, xlim = c(0,100))

# this looks okay
```
## CHL MODIFIED
```{r}
# VERY skewed
hist(hilomodified$ChlRFU, main = "Histogram of Chlorophyll", xlab = "Chlorophyll  - relative fluorescence units (RFU)")

# this looks better
hist(log(hilomodified$ChlRFU), main = "Histogram of Log Chlorophyll", xlab = "Chlorophyll")
```
## Turbitity MODIFIED
```{r}
# VERY skewed
hist(hilomodified$Corr.NTU, main = "Histogram of Turbidity", xlab = "Turbidity - Nephelometric Turbidity Units (NTU)")

# this looks better
hist(log(hilomodified$Corr.NTU), main = "Histogram of Log Turbidity", xlab = "Turbidity")

```
## Salinity MODIFIED
```{r}
# skewed
hist(hilomodified$saltppt, main = "Histogram of Salinity", xlab = "Salinity - unit parts per thousand (PPT)")

```
## Temperature MODIFIED
```{r}
# skewed
hist(hilomodified$TempC, main = "Histogram of Temperature", xlab = "Temperature - C")

```
## Dissolved Oxygen MODIFIED
```{r}
# skewed
hist(hilomodified$saltppt, main = "Histogram of Dissolved Oxygen", xlab = "Dissolved Oxygen")

```
=============================================================

# Plot with ALL Var 2013-2015
It is hard to see what is going on 
```{r}
AllYears <- ggplot(hilomodified,  aes(x = date, y = cms)) + 
  geom_line()+
  geom_line(aes(y =saltppt), color = "darkred") +
  geom_line(aes(y = Corr.NTU), color="darkgreen") +
  geom_line(aes(y=ChlRFU),color="blue")

print(AllYears + ggtitle("Hilo Bay")+labs(x="Time", y = "River Flow - cubic meters per second"))

```