---
title: "Data Cleaning & Descriptives"
author: "Brianna Cirillo & Odalys Barrientos"
output: html_notebook
---
# Libraries
```{r}
library(mosaic)
library(ggplot2)
library(readr)
```

# Matching the bouy data to the other data collected 
```{r}
getwd()
# hilo <- read_csv("~/Documents/MA321-SP21/HiloBay-Adolf/HiloBay_MA321_Spring21/NewDataframes/HiloData.csv")
# 
# head(hilo) # start date 2010-10-23 00:00:00	
# tail(hilo) # end date 2016-12-31 23:00:00	
```
# Removing columns we will not use in this project
```{r}
# hilo$date.2 <- NULL # another date column
# hilo$date.1 <- NULL # another date column
# hilo$BGARFU <- NULL # ?
# hilo$cfs <- NULL
# hilo$DOmgL <- NULL # dissolved oxygen
# hilo$PAR1 <- NULL # ?
# hilo$pH <- NULL # pH
# hilo$NTU <- NULL # a different measurement for turbitity
# hilo$DOper10 <- NULL # dissolved oxygen
# hilo$logcms <- NULL # dissolved oxygen
# hilo$chlcal <- NULL

# Columns we will use
# for some reason they were saved as characters
# hilo$cms <- as.numeric(hilo$cms)
# hilo$ChlRFU <- as.numeric(hilo$ChlRFU)
# hilo$saltppt <- as.numeric(hilo$saltppt)
# hilo$TempC <- as.numeric(hilo$TempC)
# hilo$Corr.NTU <- as.numeric(hilo$Corr.NTU)
# hilo$Doper <- as.numeric(hilo$Doper)

# saving this dataframe as a csv
#write.csv(hilo,"/Users/odalysbar/Documents/MA321-SP21/HiloBay-Adolf/HiloBay_MA321_Spring21/NewDataframes/HiloFull-6var.csv", row.names = FALSE)

# This is the dataframe we will be using 
hilo <- read_csv("~/Documents/MA321-SP21/HiloBay-Adolf/HiloBay_MA321_Spring21/NewDataframes/HiloFull-6var.csv")
hilo$cms <- as.numeric(hilo$cms)
head(hilo) # start date 2010-10-23 00:00:00	
tail(hilo) # end date 2016-12-31 23:00:00	
```

====================================================
# FULL DATA SET 2010 (Oct 23) - 2016
# Descriptives: PLots

## River Flow FULL DATA SET
```{r}
length(hilo$cms[which(is.na(hilo$cms)==TRUE)]) # 12 NAs 
which(is.na(hilo$cms)==TRUE) # the location of the NAs

RiverFlow <- ggplot(hilo,  aes(x = date, y = cms)) + 
  geom_line()

print(RiverFlow + ggtitle("River Flow")+labs(x="Time", y = "River Flow - cubic meters per second"))


favstats(hilo$cms)
```
## CHL FULL DATA SET
```{r}
length(hilo$ChlRFU[which(is.na(hilo$ChlRFU)==TRUE)]) # 20464 NAs


CHL <- ggplot(hilo,  aes(x = date, y = ChlRFU)) + 
  geom_line()

print(CHL + ggtitle("Chlorophyll ")+labs(x="Time", y = "Chlorophyll  - relative fluorescence units (RFU)"))

favstats(hilo$ChlRFU)
```
## Turbity FULL DATA SET
```{r}
length(hilo$Corr.NTU[which(is.na(hilo$Corr.NTU)==TRUE)]) #15012 NAs

TURB <- ggplot(hilo,aes(x = date, y = Corr.NTU)) + 
  geom_line()

print(TURB + ggtitle("Turbidity ")+labs(x="Time", y = "Turbidity - Nephelometric Turbidity Units (NTU)"))

favstats(hilo$Corr.NTU)
```
## Salinity FULL DATA SET
```{r}
length(hilo$saltppt[which(is.na(hilo$saltppt)==TRUE)]) #11330 NAs

SALT <- ggplot(hilo,  aes(x = date, y = saltppt)) + 
  geom_line()

print(SALT + ggtitle("Salinity")+labs(x="Time", y = "Salinity - unit parts per thousand (PPT)"))

favstats(hilo$saltppt)
```
## Tempurature FULL DATA SET
```{r}
length(hilo$TempC[which(is.na(hilo$TempC)==TRUE)]) # 10410 NAs

TEMP <- ggplot(hilo,  aes(x = date, y = TempC)) + 
  geom_line()

print(TEMP + ggtitle("Temperature")+labs(x="Time", y = "Temperature - Celsius"))

favstats(hilo$TempC)
```
# Dissolved Oxygen FULL DATA SET
```{r}
length(hilo$Doper[which(is.na(hilo$Doper)==TRUE)]) # 2267 NAs

DISSO <- ggplot(hilo,  aes(x = date, y = Doper)) + 
  geom_line()

print(DISSO + ggtitle("Dissolved Oxygen")+labs(x="Time", y = "Dissolved Oxygen in percent of saturation"))

favstats(hilo$Doper)
```

========================================================
# Histograms FULL DATA SET
## River Flow Histogram
```{r}
hist(hilo$cms, main = "Histogram of River Flow", xlab = "River Flow", breaks = 1000, ylim= c(0,400))
# VERYY SKEWED

# this looks better
hist(log(hilo$cms), main = "Histogram of Log River Flow", xlab = "Log River Flow")
```
## CHL Histogram
```{r}
# VERY skewed
hist(hilo$ChlRFU, main = "Histogram of Chlorophyll", xlab = "Chlorophyll  - relative fluorescence units (RFU)")

# this looks better
hist(log(hilo$ChlRFU), main = "Histogram of Log Chlorophyll", xlab = "Chlorophyll")

```

## Turbity Histogram
```{r}
# VERY skewed
hist(hilo$Corr.NTU, main = "Histogram of Turbidity", xlab = "Turbidity - Nephelometric Turbidity Units (NTU)")

# this looks better
hist(log(hilo$Corr.NTU), main = "Histogram of Log Turbidity", xlab = "Turbidity")

```
## Salinity Histogram
```{r}
# skewed
hist(hilo$saltppt, main = "Histogram of Salinity", xlab = "Salinity - unit parts per thousand (PPT)")

# this is worst!
hist(log(hilo$saltppt), main = "Histogram of Log Salinity", xlab = "Salinity")

```
## Temp Histogram
```{r}
# looks good
hist(hilo$TempC, main = "Histogram of Temperature", xlab = "Temperature - Celsius")
```

## Dissolved Oxygen Histogram
```{r}
# looks good 
hist(hilo$Doper, main = "Histogram of Dissolved Oxygen", xlab = "Dissolved Oxygen in percent of saturation")

```

=========================================================
# NEW DATA SET-Modified Data 2013-2015
```{r}
 
hilomodified <- read_csv("~/Documents/MA321-SP21/HiloBay-Adolf/HiloBay_MA321_Spring21/NewDataframes/HiloModified13-15.csv")

# start date: 2013-01-01 00:00:00
# end date: 2015-12-31 23:00:00

head(hilomodified)
tail(hilomodified)

```

# Descriptives on all variables MODIFIED DATA SET
## River Flow MODIFIED
```{r}
length(hilomodified$logcms[which(is.na(hilomodified$logcms)==TRUE)]) # No NAs, YAY!


RiverFlowMod <- ggplot(hilomodified,  aes(x = date, y = cms)) + 
  geom_line()

print(RiverFlowMod + ggtitle("River Flow")+labs(x="Time", y = "River Flow - cubic meters per second"))

favstats(hilomodified$cms)
```
## CHL MODIFIED
```{r}
sum(is.na(hilomodified$ChlRFU)==TRUE) # 1884 NAs

CHLMod <- ggplot(hilomodified,  aes(x = date, y = ChlRFU)) + 
  geom_line()

print(CHLMod + ggtitle("Chlorophyll ")+labs(x="Time", y = "Chlorophyll  - relative fluorescence units (RFU)"))

favstats(hilomodified$ChlRFU)
```
## Turbitity MODIFIED
```{r}
length(hilomodified$Corr.NTU[which(is.na(hilomodified$Corr.NTU)==TRUE)]) # 2704 NAs

TURBMod <- ggplot(hilomodified,  aes(x = date, y = Corr.NTU)) + 
  geom_line()

print(TURBMod + ggtitle("Turbidity ")+labs(x="Time", y = "Turbidity - Nephelometric Turbidity Units (NTU)"))

favstats(hilomodified$Corr.NTU)
```
## Salinity MODIFIED
```{r}
length(hilomodified$saltppt[which(is.na(hilomodified$saltppt)==TRUE)]) # 2267 NAs

SALTMod <- ggplot(hilomodified,  aes(x = date, y = saltppt)) + 
  geom_line()

print(SALTMod + ggtitle("Salinity")+labs(x="Time", y = "Salinity - unit parts per thousand (PPT)"))

favstats(hilomodified$saltppt)
favstats(hilomodified$TempC)
favstats(hilomodified$Doper)
```
## Temperature MODIFIED
```{r}
length(hilomodified$TempC[which(is.na(hilomodified$TempC)==TRUE)]) # 1702 NAs

TempMod <- ggplot(hilomodified,  aes(x = date, y = TempC)) + 
  geom_line()

print(TempMod + ggtitle("Temperature")+labs(x="Time", y = "Temperature - C"))


favstats(hilomodified$TempC)
```

## Dissolved Oxygen MODIFIED
```{r}
length(hilomodified$Doper[which(is.na(hilomodified$Doper)==TRUE)]) # 2267 NAs

DissoMod <- ggplot(hilomodified,  aes(x = date, y = Doper)) + 
  geom_line()

print(DissoMod + ggtitle("Dissolved Oxygen")+labs(x="Time", y = "Dissolved Oxygen"))

favstats(hilomodified$Doper)
```

===================================================
# Histograms MODIFIED 

## River Flow MODIFIED
```{r}
hist(hilomodified$cms, main = "Histogram of River Flow", xlab = "River Flow", breaks =90, xlim = c(0,100))

# this looks okay
```
## CHL MODIFIED
```{r}
# VERY skewed
hist(hilomodified$ChlRFU, main = "Histogram of Chlorophyll", xlab = "Chlorophyll  - relative fluorescence units (RFU)")

# this looks better
hist(log(hilomodified$ChlRFU), main = "Histogram of Log Chlorophyll", xlab = "Chlorophyll")
```
## Turbitity MODIFIED
```{r}
# VERY skewed
hist(hilomodified$Corr.NTU, main = "Histogram of Turbidity", xlab = "Turbidity - Nephelometric Turbidity Units (NTU)")

# this looks better
hist(log(hilomodified$Corr.NTU), main = "Histogram of Log Turbidity", xlab = "Turbidity")

```
## Salinity MODIFIED
```{r}
# skewed
hist(hilomodified$saltppt, main = "Histogram of Salinity", xlab = "Salinity - unit parts per thousand (PPT)")

```
## Temperature MODIFIED
```{r}
# skewed
hist(hilomodified$TempC, main = "Histogram of Temperature", xlab = "Temperature - C")

```
## Dissolved Oxygen MODIFIED
```{r}
# skewed
hist(hilomodified$saltppt, main = "Histogram of Dissolved Oxygen", xlab = "Dissolved Oxygen")

```
=============================================================

# Plot with ALL Var 2013-2015
It is hard to see what is going on 
```{r}
AllYears <- ggplot(hilomodified,  aes(x = date, y = cms)) + 
  geom_line()+
  geom_line(aes(y =saltppt), color = "darkred") +
  geom_line(aes(y = Corr.NTU), color="darkgreen") +
  geom_line(aes(y=ChlRFU),color="blue")

print(AllYears + ggtitle("Hilo Bay")+labs(x="Time", y = "River Flow - cubic meters per second"))

```

=========================================================

# ATTEMPT 1: Separating the Data by Storm using SALT

```{r}

outcomes <- as.numeric(hilomodified$saltppt)<25 
index.lt25 <- which(outcomes==TRUE)

poss.storms <- hilomodified[index.lt25,]
poss.storms
```


```{r}
outcomes.35 <- as.numeric(hilomodified$saltppt)<35 
index.lt35 <- which(outcomes==TRUE)
index.lt25
poss.storms35 <- hilomodified[index.lt25,]
poss.storms35

hilomodified$date[16]
```

```{r}
storm.1.7.13 <- ggplot(hilomodified[145:168,],  aes(x = date, y = as.numeric(cms))) + 
  geom_line(color="black")+
  geom_line(aes(y = as.numeric(saltppt)), color = "lightsteelblue1") +
  geom_line(aes(y = as.numeric(Corr.NTU)), color="grey69") +
  geom_line(aes(y=as.numeric(ChlRFU)),color="khaki")

print(storm.1.7.13 + ggtitle("Storm 1/7/13")+labs(x="Time"))
```

# ATTEMPT 2: Separating the Data by Storm using CHANGE IN RIVER FLOW

```{r}
# any value greater than or equal to 10 cms indicates a storm
pos.neg <- as.numeric(hilomodified$cms) - 10 

start <- vector()
end <- vector()

for(i in 1:length(pos.neg)){
  if(isTRUE(pos.neg[i] < 0 && pos.neg[i+1] > 0)){ 
    # if goes from + to - then this is the beginning of a storm
    start[i] <- i
  }else if(isTRUE(pos.neg[i] > 0 && pos.neg[i+1] < 0)){
    # if goes from - to + then this is the end of the storm
    end[i] <- i
    }
}
start<-start[!is.na(start)] # removing NA
start
length(start) # make sure both start and end are the same length 

end<-end[!is.na(end)]
end
length(end)
```

```{r}
for(i in 1:length(start)){
  # the point start[i] is the beginning of a storm
  # the point end[i] is the end of the same storm
  # we are adding 24 hours to each start and end time
 p <- ggplot(hilomodified[(start[i]-24):(end[i]+24),],aes(x = date, y = as.numeric(cms))) +
  geom_line(color="black")
 p <- p + geom_line(aes(y = range01(as.numeric(Corr.NTU))), color="green") 
 p <- p + geom_line(aes(y= range01(as.numeric(ChlRFU))),color="blue") 
 p <- p + scale_y_continuous(name = "First Axis", sec.axis = sec_axis(trans = ~.*.05, name="Second Axis")
  ) 
  # geom_line(aes(y = range01(as.numeric(Corr.NTU))), color="green") +
  # geom_line(aes(y= range01(as.numeric(ChlRFU))),color="blue") +
  # geom_line(aes(y = range01(as.numeric(saltppt))),color="red") +
  # geom_line(aes(y = range01(as.numeric(TempC))),color="purple") +
  # geom_line(aes(y = range01(as.numeric(Doper))),color="orange")+
   p <- p + ggtitle("Storms")+labs(x="Time")
 print(p)
}

```

```{r}
range01 <- function(x, na.rm = TRUE) {
  # version 1.1 (9 Aug 2013)
  (x - min(x, na.rm = na.rm)) / (max(x, na.rm = na.rm) - min(x, na.rm = na.rm))
}
```

# ATTEMP 3: Separating the Data by Storm RIVER FLOW > 10cms
WE WILL USE THIS METHOD
# using base R
```{r,fig.height=20}
#num <- c(1:length(start))
#graphics.off()

for(i in 1:length(start)){
#for(i in c(1)){
  
#jpeg(sprintf("storm-%s.jpeg", num[i]),width=2000, height = 1000)
  
index <- c(start[i]:(end[i]+24))

date <- c(hilomodified$date[index])
enddate <- hilomodified$date[end[i]+24]-10
endcms <- as.numeric(hilomodified$cms[end[i]+24])
endchl <- as.numeric(hilomodified$ChlRFU[end[i]+24])
endturb <- as.numeric(hilomodified$Corr.NTU[end[i]+24])
endsalt <- as.numeric(hilomodified$saltppt[end[i]+24])
endtemp <- as.numeric(hilomodified$TempC[end[i]+24])
endDO <- as.numeric(hilomodified$Doper[end[i]+24])


par(mfrow=c(6,1)) 
# RIVER FLOW
plot(x = date,y = as.numeric(hilomodified$cms[index]), type = "l",lwd=2, col="black", xlab="Time", ylab="River Flow cms", cex.lab = 1.5) # Create first plot for River Flow
text(x= enddate, y = endcms, col = "black", adj=1, cex=3, label= "River Flow")

# CHLOROPHYLL
if(sum(is.na(as.numeric(hilomodified$ChlRFU[index])))==length(hilomodified$ChlRFU[index])){
  plot(x = 0 ,y = 0, type = "l",lwd=2, col="blue", xlab= "Time", ylab = "Chlorophyll  - relative fluorescence units (RFU)", cex.lab = 1.5)
text(x = 0, y = 0, col = "blue", adj=1, cex=3, label= "Chlorophyll")
}else{
plot(x = date ,y = as.numeric(hilomodified$ChlRFU[index]), type = "l",lwd=2, col="blue", xlab= "Time", ylab = "Chlorophyll  - relative fluorescence units (RFU)", cex.lab = 1.5)
text(x = enddate, y = endchl, col = "blue", adj=1, cex=3, label= "Chlorophyll")
}

# TURBIDITY
if(sum(is.na(as.numeric(hilomodified$Corr.NTU[index])))==length(hilomodified$Corr.NTU[index])){
  plot(x = 0 ,y = 0, type = "l",lwd=2, col="red", xlab= "Time", ylab = "Turbidity - Nephelometric Turbidity Units (NTU)", cex.lab = 1.5)
text(x = 0, y = 0, col = "red", adj=1, cex=3, label= "Turbidity")
}else{
plot(x = date ,y = as.numeric(hilomodified$Corr.NTU[index]), type = "l",lwd=2, col="red", xlab= "Time", ylab = "Turbidity - Nephelometric Turbidity Units (NTU)", cex.lab = 1.5)
text(x = enddate, y = endturb, col = "red", adj=1, cex=3, label= "Turbidity")
}

# SALINITY
if(sum(is.na(as.numeric(hilomodified$saltppt[index])))==length(hilomodified$saltppt[index])){
  plot(x = 0 ,y = 0, type = "l",lwd=2, col="green", xlab= "Time", ylab = "Salinity - unit parts per thousand (PPT)", cex.lab = 1.5)
text(x = 0, y = 0, col = "green", adj=1, cex=3, label= "Salinity")
}else{
plot(x = date ,y = as.numeric(hilomodified$saltppt[index]), type = "l",lwd=2, col="green", xlab= "Time", ylab = "Salinity - unit parts per thousand (PPT)", cex.lab = 1.5)
text(x = enddate, y = endsalt, col = "green", adj=1, cex=3, label= "Salinity")
}

# TEMPERATURE
if(sum(is.na(as.numeric(hilomodified$TempC[index])))==length(hilomodified$TempC[index])){
  plot(x = 0 ,y = 0, type = "l",lwd=2, col="orange", xlab= "Time", ylab = "Temperature - Celsius", cex.lab = 1.5)
text(x = 0, y = 0, col = "orange", adj=1, cex=3, label= "Temperature")
}else{
plot(x = date ,y = as.numeric(hilomodified$TempC[index]), type = "l",lwd=2, col="orange", xlab= "Time", ylab = "Temperature - Celsius", cex.lab = 1.5)
text(x = enddate, y = endtemp, col = "orange", adj=1, cex=3, label= "Temperature")
}

# DISSOLVED OXYGEN
if(sum(is.na(as.numeric(hilomodified$Doper[index])))==length(hilomodified$Doper[index])){
  plot(x = 0 ,y = 0, type = "l",lwd=2, col="purple", xlab= "Time", ylab = "Dissolved Oxygen in percent of saturation", cex.lab = 1.5)
text(x = 0, y = 0, col = "purple", adj=1, cex=3, label= "Dissolved Oxygen")
}else{
plot(x = date ,y = as.numeric(hilomodified$Doper[index]), type = "l",lwd=2, col="purple", xlab= "Time", ylab = "Dissolved Oxygen in percent of saturation", cex.lab = 1.5)
text(x = enddate, y = endDO, col = "purple", adj=1, cex=3, label= "Dissolved Oxygen")
}

#dev.off()
}

```
 



# ============================= Dickey Fuller Test ===============================

```{r}
library(tseries)
adf.test(hilo2015.fixed$cms, alternative= "stationary")
adf.test(hilo2015.fixed$ChlRFU, alternative= "stationary")
adf.test(hilo2015.fixed$Doper, alternative= "stationary")
adf.test(hilo2015.fixed$saltppt, alternative= "stationary")
adf.test(hilo2015.fixed$TempC, alternative= "stationary")
adf.test(hilo2015.fixed$Corr.NTU, alternative= "stationary")
```

 
















