---
title: "Yearly Descriptives"
author: "Brianna Cirillo & Odalys Barrientos"
output: html_notebook
---
# Data starts from HiloModified13-15.csv
```{r}
library(mosaic)
library(ggplot2)
library(readr)

hilomodified <- read_csv("~/Documents/MA321-SP21/HiloBay-Adolf/HiloBay_MA321_Spring21/NewDataframes/HiloModified13-15.csv")
```

==============================================================
# Descriptives by Storm
We are picking one storm from each year. We can indicate a storm has occurred by the extreme events in the river flow data. 
We will not use the log (which is logbase10) in order to see the extreme events
When salinity is below 35 this also indicates a storm has occurred. 

We will break the data set by year to find the most extreme event for each year. 
```{r}
hilo2013 <- hilomodified[1:8760,]

hilo2014 <- hilomodified[8761:17520,]

hilo2015 <- hilomodified[17521:length(hilomodified[,1]),]
# We will save hilo2015 as a csv file for future use
# write.csv(hilo2015,"/Users/odalysbar/Documents/MA321-SP21/HiloBay-Adolf/HiloBay_MA321_Spring21/NewDataframes/Hilo2015.csv", row.names = FALSE)

```

# 2013 Data & Plot
```{r}
# max(as.numeric(hilo2013$logcms)) # this is log base 10 
# This is NOT the natural log
max(hilo2013$cms) # 207.186 
which(hilo2013$cms == max(hilo2013$cms))

hilo2013[3550,]
# all values for this storm are NA

plot2013 <- ggplot(hilo2013,  aes(x = date, y = cms)) + 
  geom_line(color="black")+
  geom_line(aes(y = saltppt), color = "darkred") +
  geom_line(aes(y = Corr.NTU), color="darkgreen") +
  geom_line(aes(y=ChlRFU),color="blue")

print(plot2013 + ggtitle("2013")+labs(x="Time", y = "River Flow - cubic meters per second"))
```

#### Split 2013 into 6 months to get a better visual
```{r}
# R2013.1 <- ggplot(hilo2013[1:(length(hilo2013[,1])/2),],  aes(x = date, y = as.numeric(cms))) + 
#   geom_line(color="black")+
#   geom_line(aes(y = as.numeric(saltppt)), color = "darkred") +
#   geom_line(aes(y = as.numeric(Corr.NTU)), color="darkgreen") +
#   geom_line(aes(y=as.numeric(ChlRFU)),color="blue")
# print(R2013.1 + ggtitle("River Flow")+labs(x="Time", y = "River Flow - cubic meters per second"))
# R2013.1+ylim(0,40)
# 
# 
# R2013.2 <- ggplot(hilo2013[(length(hilo2013[,1])/2):length(hilo2013[,1]),],  aes(x = date, y = as.numeric(cms))) + 
#   geom_line(color="black")+
#   geom_line(aes(y = as.numeric(saltppt)), color = "darkred") +
#   geom_line(aes(y = as.numeric(Corr.NTU)), color="darkgreen") +
#   geom_line(aes(y=as.numeric(ChlRFU)),color="blue")
# print(R2013.2 + ggtitle("River Flow")+labs(x="Time", y = "River Flow - cubic meters per second"))
# R2013.2+ylim(0,40)
```

# 2014 Data & Plot
```{r}
# Max river flow in the overall data set
max(hilo2014$cms)
max(hilomodified$cms)

which(hilo2014$cms == max(hilo2014$cms))

hilo2014[5263,]

R2014 <- ggplot(hilo2014,  aes(x = date, y = cms)) + 
  geom_line(color="black")+
  geom_line(aes(y = saltppt), color = "darkred") +
  geom_line(aes(y = Corr.NTU), color="darkgreen") +
  geom_line(aes(y=ChlRFU),color="blue")

print(R2014 + ggtitle("2014")+labs(x="Time", y = "River Flow - cubic meters per second"))
```


# 2015 Data & Plot
```{r}
max(hilo2015$cms)

which(hilo2015$cms == max(hilo2015$cms))

hilo2015[6475,]

R2015 <- ggplot(hilo2015,  aes(x = date, y = cms)) + 
  geom_line(color="black")+
  geom_line(aes(y = saltppt), color = "darkred") +
  geom_line(aes(y = Corr.NTU), color="darkgreen") +
  geom_line(aes(y=ChlRFU),color="blue")

print(R2015 + ggtitle("2015")+labs(x="Time", y = "River Flow - cubic meters per second"))
```
























